CREATING AN DIRECTED SOCIAL NETWORK FROM 'agents.csv'
===
based on the empirical homophily at Holzhauer
---

LOADING AGENTS
```{r}

setwd("/Users/thorbenje/phd/models/chapter4/input/")

# load households
hh <- read.csv("agents.csv", sep=" ")
hh <- hh[,2:5]

# change row names
# row.names(hh) <- as.numeric(row.names(hh)) - 1

```



CALCULATE DISTANCE MATRIX (UP TO ca. 140 METERS) AND COMPRESS DISTANCE MATRIX TO 'reachable_neighbors' ARRAY
```{r}
library(FNN)

hh_coord <- cbind( hh$x, hh$y )
nn <- get.knn(hh_coord, k=999)

# change row names
row.names(nn) <- as.numeric(row.names(nn)) - 1

## how matrix connects data frames:
## nn$nn.index: household ID + 1 = row; neighbor ID + 1 = column

# parameter
d_NBHD <- 20   # max. distance between nodes that are 'within the same neighborhood'

potential_neighbors <- list()

# iterate over all rows of nn
for (i in 1:nrow(nn$nn.index)) {
  
  # select those elements of the row that are within 'd_NBHD'
  number_potential_neighbors <- length( which( nn$nn.dist[i,] < d_NBHD) )

  # append those elements as new sublist to list
  potential_neighbors[i] <- list(nn$nn.index[i, 1:number_potential_neighbors ])  
}

# remove 'nn' from RAM
rm(nn)

# HOWTO retrieve IDs of nearest neighbors for household i: ' unlist(potential_neighbors[i+1])-1 '
```

Homophily function: Calculate in_link based on homophily
```{r}

# defining probability matrix
homophily_matrix <- matrix(c(.59, .10, .22, .09, .50, .34, .12, .04, .36, .25, .32, .07, .37, .15, .32, .16), ncol=4, byrow=T)
homophily_matrix_tags <- c("GLM", "TRA", "MAI", "HED")

# clustering of lifestyles to make it compatible with Holzhauer
lifestyle <- unique(as.vector(hh$group))
lifestyle_cluster <- matrix( c(lifestyle, c("GLM", "TRA", "GLM", "MAI", "HED")), ncol=2 )


homophily <- function( lifestyle, number ) {
  # lifestyle group belonging to
  cluster <- lifestyle_cluster[ lifestyle_cluster[,1] == lifestyle , 2 ] 
  # lifestyle cluster
  matrix_index <- which(homophily_matrix_tags == cluster)
  # draw 'number' other lifestyles to connect to
  sample <- sample(1:4, size = number, replace = TRUE, prob = homophily_matrix[matrix_index,])
  
  return (homophily_matrix_tags[sample])
}

cluster <- function( lifestyle ) {
  return( lifestyle_cluster[ lifestyle == lifestyle_cluster[,1], 2] )
}

# VALIDATION: print histogram of connections for lifestyles
# barplot(prop.table(table( homophily("leader", 10000) )))

```

NETWORK GENERATION STEP 1/3 - LOCAL CONNECTIONS
```{r}
library(rgeos)

# initialize degree and deg_target
degrees <- c(1, 2, 2, 2, 3, 3, 4, 5, 6, 7, 8)   # from SusLab interviews
random_degrees <- sample(degrees, nrow(hh), replace=T)
hh$deg_target <- random_degrees
hh$degree <- 0

# copy relevant content from 'hh'
households <- as.matrix(data.frame(ID=as.numeric(hh$id), 
                                   degree=as.numeric(hh$degree), 
                                   deg_target=as.numeric(hh$deg_target)))

# store edges globally ([1] = id1; [2] = id2)
needed_edges <- sum(households[,3])

# edges (format: in - out)
edges <- matrix(data = 0, nrow = needed_edges, ncol = 2)
# where to write next edge
edge_index <- 1

## make local connections (according to degree target and d_NBHD)
# iterate over households
system.time(

  for (id1 in 1:nrow(households)) {
#   for (id1 in 0:1000) {

    # print( paste("Assigning neighbors to household",id1,"of",length(hh$ID),"households..."))
    if (id1 %% 100 == 0) print( paste("assigning local neighbors to household", paste(id1), "of", nrow(households)))
    
    # calculate needed connections for this node
    needed_connetions <- households[id1,3] - households[id1,2]
    # IF no needed connections OR greater degree than deg_target: skip id
    if( needed_connetions <= 0 ) next

    free_neighbors <- potential_neighbors[[id1]] # get IDs of free neighbors (index correction)
    free_neighbors <- free_neighbors[ ! free_neighbors %in% edges[ edges[,1] == id1, 2] ] # remove those FROM which already a connection exists 
    free_neighbors_cluster <- unlist( lapply( hh$group[free_neighbors], cluster) )   #list 'lifestyle cluster' of these free neighbors

    ## consider homophily
    # Draw set of lifestyles to connect to (based on Holzhauser's empirical homophily)
    neighbors_lifestyles <- homophily( hh$group[id1], needed_connetions )

    # For each drawn lifestyle, sample a free_neighbor with that lifestyle - result: 'new neighbors'
    new_neighbors <- numeric()
    for (c in unique(neighbors_lifestyles)) {
      
      # number of times this cluster occurs
      frequence <- as.numeric( table(neighbors_lifestyles)[c] )
      
      # select free_neighbors with from this 'cluster'
      free_neighbors_from_cluster <- free_neighbors[ which(free_neighbors_cluster == c) ]
      
      # IF no free neighbors with this lifestyle: leave for loop
      if (length(free_neighbors_from_cluster) == 0) next
      # IF more free neighbors than needed: lower 'frequence'
      if (frequence > length(free_neighbors_from_cluster)) frequence <- length(free_neighbors_from_cluster)
      
      # sample 'frequence' from free_neighbors  - BUG: sampling not from free_neighbors_from_cluster??
      if (length(free_neighbors_from_cluster) > 1) { 
        chosen_neighbors <- sample(free_neighbors_from_cluster, frequence, replace = FALSE)
      } else {
        chosen_neighbors <- free_neighbors_from_cluster
      }
      
      # ASSERT: chosen neighbors within d_NBHD of id1
      if( Reduce("&", ! chosen_neighbors %in% unlist(potential_neighbors[id1])) ) stop( paste("Id1 ",id1,"NOT within d_NBHD of id2",chosen_neighbors))
      
      # add to 'new_neighbors'
      new_neighbors <- c(new_neighbors, chosen_neighbors)
    }
    
    # VALIDATE lifestyle of new_neighbors:
    # sort(unlist( lapply( lifestyle[new_neighbors], cluster) ))

    # IF no free neighbors: skip id1
    if ( length(new_neighbors) == 0 ) next
    
    # assign new neighbors to household
    for (id2 in new_neighbors) {
      
      edges[ edge_index, ] <- c(id1,id2)
      edge_index <- edge_index + 1
    }
    
    # update in-degree for id1 in hh
    households[id1,2] <- households[id1,2] + length( new_neighbors )
  }
)

## GRAPH QUALITY ASSESSMENT
# ASSERT: no household with greater degree than deg_target
households_exceeding_target <- nrow( households[ households[,2] > households[,3] ,])
if (households_exceeding_target > 0) warning(paste( households_exceeding_target, "households with greater degree than deg_target!"))

# ASSERT: no household with no links
households_without_degree <- nrow( households[ households[,2] == 0 ,])
if (households_without_degree > 0) {
  warning(paste( households_without_degree, "households with no local links!"))
  warning(paste(round((length( edges[ edges[,1] != 0 ,1] ) / length( edges[,1] )) * 100, 2), "% of local edges created"))
}

```

NETWORK GENERATION STEP 2/3 - REMOVING SOME EDGES
```{r}

# copy data
households2 <- households
edges2 <- edges

# parameters
p_NBHD <- 0.5   # probability of a link to be 'within the neighborhood'

## remove some local connections (according to p_NBHD parameter)
# randomly select edges that shall be removed
indexes_to_remove <- sort( sample( 1:nrow(edges2), round( p_NBHD * nrow(edges2)) ) )

## update in-degree of households
# count times a household is connected to an edge that is to be removed (only considering in-degree)
deg_decrement <- data.frame( table( edges2[indexes_to_remove,1]) )
names(deg_decrement) <- c("ID", "decrement")

# iterate over households whose edges were removed
for (id in deg_decrement$ID) {
  
  id_numeric <- as.numeric(id)
  if( id_numeric %% 100 == 0) print(paste("Decreasing in-degree for node ", id, "of max. 32.000"))
  
  # in households, reduce degree
  households2[id_numeric,2] <- households2[id_numeric,2] - deg_decrement[deg_decrement$ID == id,]$decrement
}

# ASSERT: no hosehold in 'households' has negative degree, now
if( nrow( households2[ households2[,2] < 0 ,]) > 0) stop( "households2 contains households with negative degree!" )

# remove lines
edges2 <- edges2[ - indexes_to_remove ,]
```

NETWORK GENERATION STEP 3/3 - GLOBAL CONNECTIONS
```{r}

# copy edge vector and households from above
households3 <- households2
edges3 <- edges2

# remove potential zero lines from edges3
edges3 <- edges3[edges3[,1] != 0,]

# calculate size of matrix to be allocated
total_edges <- sum(households3[,3])
rows_now <- nrow(edges3)
rows_needed <- total_edges - rows_now

# allocating new 'edge3' matrix with empty rows
zero_vector <- rep(0, rows_needed)
edges3 <- cbind( c(edges3[,1], zero_vector), c(edges3[,2], zero_vector))
# initialize writing index
edge_index <- rows_now + 1

## 'fill up' with global connections
system.time(
  
  for (id1 in 1:nrow(households)) {
    
    if (id1 %% 100 == 0) print( paste("assigning global neighbors to household", paste(id1), "of 31840"))
    
    # calculate needed connections for this node
    needed_connections <- households3[id1,3] - households3[id1,2]
    # IF no needed connections OR greater degree than deg_target: skip id
    if( needed_connections <= 0 ) next
    # IF needed conenction negative: report error
#     if( needed_connetions < 0 ) stop ("Household with ID=id1 has greater degree than deg_target!")
    
    # get IDs of other households
    other_households <- households3[ households3[,1] != id1, 1 ]
    # remove those households from 'other_households' FROM which already a connection exists (in-link to id1)
    other_households <- other_households[ ! other_households %in% edges3[ edges3[,1] == id1, 2] ]
    # list 'lifestyle cluster' of these free neighbors
    other_households_cluster <- unlist( lapply( hh$group[other_households], cluster) )

    ## consider homophily
    # Draw set of lifestyles to connect to (based on Holzhauser's empirical homophily)
    neighbors_lifestyles <- homophily( hh$group[id1], needed_connections )

    # For each drawn lifestyle, sample a free_neighbor with that lifestyle - result: 'new neighbors'
    new_neighbors <- numeric()
    for (c in unique(neighbors_lifestyles)) {
      
      # print(tag)
      # number of times this cluster occurs
      frequence <- as.numeric( table(neighbors_lifestyles)[c] )
      
      # select free_neighbors with from this 'cluster'
      free_households_from_cluster <- other_households[ which(other_households_cluster == c) ]
      # PERFORMANCE OPTION: remove here: id1 and households from which a link exists
      
      # IF no free neighbors with this lifestyle: leave for loop
      if (length(free_households_from_cluster) == 0) next
      # IF more free neighbors than needed: lower 'frequence'
      if (frequence > length(free_households_from_cluster)) frequence <- length(free_households_from_cluster)
      
      # sample 'frequence' from free_neighbors  - BUG: sampling not from free_neighbors_from_cluster??
      if (length(free_households_from_cluster) > 1) { 
        chosen_neighbors <- sample(free_households_from_cluster, frequence, replace = FALSE)
      } else {
        chosen_neighbors <- free_households_from_cluster
      }
      
      # add to 'new_neighbors'
      new_neighbors <- c(new_neighbors, chosen_neighbors)
    }
    
    # VALIDATE lifestyle of new_neighbors:
    # sort(unlist( lapply( lifestyle[new_neighbors], cluster) ))

    # IF no free neighbors: skip id1
    if ( length(new_neighbors) == 0 ) next
    
    # assign new neighbors to household
    for (id2 in new_neighbors) {
      
      edges3[ edge_index, ] <- c(id1,id2)
      edge_index <- edge_index + 1
    }
    
    # update in-degree for id1 in hh
    households3[id1,2] <- households3[id1,2] + length( new_neighbors )
  }
)

## RESULT QUALITY
# TESET: no household with greater degree than deg_target
households_deg_exceeding <- length( households3[ households3[,2] > households3[,3] ,1] )
if ( households_deg_exceeding > 0) warning( paste(households_deg_exceeding,"household(s) with greater degree than deg_target!") )
# TEST: every household has enough links
households_deg_lower <- length( households3[ households3[,2] < households3[,3] ,1] )
if ( households_deg_lower > 0) warning( paste(households_deg_lower,"household(s) with lower degree than deg_target!") )
# ASSERT: every household has at least one link
households_deg_lower <- length( households3[ households3[,2] == 0 ,1] )
if ( households_deg_lower > 0) stop ( paste(households_deg_lower,"household(s) are not connected to any edges!") )


# CORRECTIONS
# sort data.frame 'edges3' by 1st column:
edges_final           <- as.data.frame(edges3)
colnames(edges_final) <- c("to", "from")
edges_final           <- edges_final[order(edges_final$to),]

# correct index of 'to' column in edges3
edges_final$to <- edges_final$to - 1

# remove self-links
edges_final <- edges_final[ edges_final$from != edges_final$to,]
# edges_final[edges_final$from == edges_final$to,]

# write edges3 to file
write.table(edges_final, "network.csv", sep="\t", row.names = FALSE)
```